Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 trunk/paper_enahpc14/Sec4:Experimental-results.tex
Manuel's changes

\subsection{Power-performance tracing of \cosmoart}
\label{subsec:4.3}

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution. First we run \cosmoart for a full-day simulation
and analyze the power consumption. Second we correlate performance with
power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23-24h).

@@

In order to obtain an initial overview of the power dissipation of the system model, Figure~\ref{fig:9} depicts the power profiles of a full-day simulation of \cosmoart leveraging the  polling and blocking MPI policies on the top and bottom plots, respectively. The power profiles shown account for the sum of dissipated power of the 16 nodes used in \tinto, captured with our PDUs. The first observation for the polling MPI policy is the quite plain power profile it generates. However, the first and last hours of the day along a 24h simulation, dissipate slightly less power than the hours than the mid-day hours. We relate these variations due to the chemical reactions and aerosols, computed by ART, that are taking place during the daylight hours but not during night. In this case peak power dissipation is about 3821\,W.
Looking at the power profile using the blocking MPI policy, we can easily observe
that the pattern is repeated favoring the reduction of power during nighttime hours in a more noticeable way. The peak power dissipation in this mode is 3768\,W, which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak power. Also, we observe power drops in each hour simulated, basically related to periods in which the processes were performing blocking-waits instead of polling-waits, thus yielding the CPU to a low usage during this phases. Our final observation is the reduction in execution time when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that the costs of the interruptions used in the blocking can lead to an increase of the total execution time, however, applications with extensive computing may show better perforamnce with the blocking mode. This plays in favour for our results since the energy consumed is also less by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

To gain insights about

>>>>

\subsection{Power-performance tracing on \tintorrum}
\label{subsec:4.3}

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution. First we run \cosmoart for a full-day simulation
and analyze the power consumption. Second we correlate performance with
power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

@@

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. o, captured with our PDUs. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case peak power 
dissipation is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

To gain more insights about the power-performance behavior of \cosmoart, we also have conducted 
experiments that show power-performance traces of 1 hour simulation. We limited the simulation 
time to this bound since the pattern of the computation performed by \cosmoart is repeated each 
hour, also because the weight of a full-day simulation trace can not be easily handled using 
our machines and memory available (about 45\,GB per trace file). To soften the problem we 
shrink the simulations to 1 hour during mid-day (11h-12h) and midnight (23h-24h), in which 
\cosmoart perform different internal computations, as shown in Figure~\ref{fig:9}. At the same
time we also leverage the two stated MPI policies: polling and waiting modes.
